---
title             : "Lab10 Re-Analysis"

author: 
  - name          : "Drew Shives"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "my@email.com"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - Conceptualization
      - Writing - Original Draft Preparation
      - Writing - Review & Editing

affiliation:
  - id            : "1"
    institution   : "The Graduate Center, CUNY"
    
header-includes:
  - \raggedbottom

output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
library("papaja")
r_refs("r-references.bib")
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

```{r, include = FALSE}
library(tidyverse)
library(pwr)
```

```{r, include = FALSE}
df <- read_csv("https://raw.githubusercontent.com/CrumpLab/statisticsLab/master/data/stroop_stand.csv")
```

```{r, include = FALSE}
head(df)
```

```{r, include = FALSE}
colMeans(df)
```

```{r, include = FALSE}
RTs <- c(as.numeric(unlist(df[,1])),
         as.numeric(unlist(df[,2])),
         as.numeric(unlist(df[,3])),
         as.numeric(unlist(df[,4]))
         )

Congruency <- rep(rep(c("Congruent","Incongruent"),each=50),2)
Posture <- rep(c("Stand","Sit"),each=100)
Subject <- rep(1:50,4)

df_long <- data.frame(Subject,Congruency,Posture,RTs)
```

```{r, include = FALSE}
df_long
```

```{r, include = FALSE}
con_stand <- df_long %>%
  filter(Congruency == "Congruent", Posture == "Stand")

incon_stand <- df_long %>%
  filter(Congruency == "Incongruent", Posture == "Stand")

con_sit <- df_long %>%
  filter(Congruency == "Congruent", Posture == "Sit")

incon_sit <- df_long %>%
  filter(Congruency == "Incongruent", Posture == "Sit")

mean(con_stand$RTs) - mean(incon_stand$RTs)

mean(con_sit$RTs) - mean(incon_sit$RTs)
```

```{r, include = FALSE}
stand_t <- t.test(con_stand$RTs, incon_stand$RTs, paired = TRUE)
```

```{r, include = FALSE}
sit_t <- t.test(con_sit$RTs, incon_sit$RTs, paired = TRUE)
```

```{r, include = FALSE}
sit_diff <- con_stand$RTs - incon_stand$RTs

stand_diff <- con_sit$RTs - incon_sit$RTs
```

```{r, include = FALSE}
sit_stand_t <- t.test(sit_diff, stand_diff, paired = TRUE)

sit_stand_t
```

```{r, include = FALSE}
grouped_ <- df_long %>%
  group_by(Congruency, Posture) %>%
  summarise(mean = mean(RTs), 
            SEM = sd(RTs) / sqrt(length(RTs)))
```

```{r, include = FALSE}
grouped_hist <- ggplot(data = grouped_, mapping = aes(x = Posture, y = mean, group = Congruency, fill = Congruency)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(mapping = aes(ymin = mean - SEM, ymax = mean + SEM),
                position = position_dodge(width = 0.9),
                width = 0.2) +
  theme_classic() +
  coord_cartesian(ylim = c(700, 1000))

```

```{r, include = FALSE}
effect_sizes <- seq(.1,2,.1)

power <- sapply(effect_sizes, FUN = function(x) {
  pwr.t.test(n=50, 
             d=x, 
             sig.level=.05, 
             type="one.sample", 
             alternative="two.sided"
             )$power
  })

plot_df <- data.frame(effect_sizes,power)

power_line <- ggplot(plot_df, aes(x=effect_sizes, y=power))+  
  geom_point()+  
  geom_line()
```

## Data analysis

We used `r cite_r("r-references.bib")` for all our analyses.

First, I was able to replicate the Fig. 1 from the original paper.

```{r, echo = FALSE}
knitr::kable(grouped_)
plot(grouped_hist)
```

Second, I was able to replicate the Stroop effects in both the sitting (`r apa_print(sit_t)$full_result`) and standing (`r apa_print(stand_t)$full_result`) conditions.

The authors found that the Stroop effect shrunk when participants were standing compared to when they were sitting, F(1, 49) = 8.964, p = .004, np2 = .155. I was able to replicate that results using a paired sample t-test, `r apa_print(sit_stand_t)$full_result`.

Third, I was able to create a power curve analysis for the design

```{r}
plot(power_line)
```


\newpage

# References

Rosenbaum, D., Mama, Y., & Algom, D. (2017). Stand by Your Stroop: Standing Up Enhances Selective Attention and Cognitive Control. Psychological science, 28(12), 1864-1867.

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id="refs" custom-style="Bibliography"></div>
\endgroup
